---
title: "Growth curve analysis using the Tecan 96-well plate reader"
author: "Saaz Sakrikar, modifying code from Grace Avecilla and David Gresham"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
---

The data comprise 288 OD measurements that were acquired over a 24 hour period.

# Growth curves requirements


* OD data from the Tecan in a csv with each well's data in a column, the first row contained the well names, and the first column is time in seconds 

* A sample sheet, with the four required columns: __Well__, __Sample__,	__Bio_rep__,	__Tech_rep__
  + The Well column did not have 0s in the well names (i.e. A1, not A01)
  + For blank wells, we put 'B' in the Sample column of the sample sheet
  + For blank wells, we put 'NA' for all other column values
  + Additional columns are possible. In this case, "Strain", "Media" (SCGlu or YPGly), "FLC_conc", "Temperature".



```{r input, warning=FALSE,message=FALSE}
library(tidyverse)
library(growthcurver)

#theme_set(theme_minimal())

cycles = 288
samplesheet = read_csv("July20_meta.csv")
#knitr::kable(samplesheet)

data_od = read_csv("July23_72hrsgrowth.csv",skip = 0, n_max = cycles) 
colnames(data_od)[1] <- "time"
#convert time to hrs from seconds
data_od$time <- as.numeric(data_od$time)/60/60

#knitr::kable(data_od)


# output csv name, used at the end only.
out = 'output.csv'

```

Transformation of data for plotting

```{r setup}
data_od_for_plot <- data_od %>% 
  pivot_longer(cols = -time, names_to = "well", values_to = "od") %>% 
  left_join(samplesheet, by = c("well" = "Well")) %>%
  unite(Sample_Well, well, Sample, sep = " ", remove = FALSE)

```

## Individual Growth Profiles

We  plot the entire dataset arrayed in 96 well format it appears on the Tecan for visual inspection.  This allows us to easily identify any wells that should be excluded.

```{r all_well_plot}
data_od_for_plot %>%
  mutate(Sample_Well = str_replace(Sample_Well, "(?<![0-9])([0-9])(?![0-9])", "0\\1")) %>%
  ggplot(aes(time, od, color = as.factor(Media))) +
  geom_line() +
  facet_wrap(~Sample_Well, ncol=12) +
  xlab("Time (hrs)") + scale_y_continuous(breaks=c(0, 2)) + scale_x_continuous(breaks=c(0, 12, 24, 36, 48, 60, 72)) + theme_bw()

```

```{r}
data_od_for_plot %>%
  filter(time <= 24) %>%
  mutate(Sample_Well = str_replace(Sample_Well, "(?<![0-9])([0-9])(?![0-9])", "0\\1")) %>%
  ggplot(aes(time, od, color = as.factor(Media))) +
  geom_line() +
  facet_wrap(~Sample_Well, ncol=12) +
  xlab("Time (hrs)") + scale_y_continuous(breaks=c(0, 2)) + scale_x_continuous(breaks=c(0, 12, 24, 36, 48, 60, 72)) + theme_bw()
```

#Obtain growth paramaters using logisitc curve

```{r analysis}
gc_out <- SummarizeGrowthByPlate(data_od, 
                                 plot_fit = FALSE) %>% 
  left_join(samplesheet, by = c("sample" = "Well")) %>%
  rename(well = sample, sample = Sample)

#knitr::kable(gc_out)

gc_sampleout=gc_out[!is.na(gc_out$Strain),]
write.csv(gc_sampleout,paste(out))
```

Plot the growth rate r and area under the curve AUC

```{r plotting}

ggplot(gc_sampleout,aes(x=Media,y=r)) + geom_dotplot(binaxis="y", stackdir="center", dotsize=1.5, aes(fill=Strain, color=Strain)) + facet_wrap(~FLC_conc) + theme_bw() + ggtitle("Growth rate r at different FLC concs")

#ggplot(gc_sampleout,aes(x=Type,y=t_gen)) + geom_dotplot(binaxis="y", stackdir="center", dotsize=2, aes(fill=Strain, color=Strain)) + facet_wrap(~FLC_conc) + theme_bw() + ggtitle("Doubling time tgen at different FLC concs")

ggplot(gc_sampleout,aes(x=Media,y=auc_l)) + geom_dotplot(binaxis="y", stackdir="center", dotsize=1.5, aes(fill=Strain, color=Strain)) + facet_wrap(~FLC_conc) + theme_bw() + ggtitle("Area under the curve AUC at diff FLC concs")

SCGluresults=gc_sampleout[gc_sampleout$Media=="SCGlu",]
YPGlyresults=gc_sampleout[gc_sampleout$Media=="YPGly",]

ggplot(SCGluresults,aes(x=as.factor(FLC_conc),y=auc_l)) + geom_dotplot(binaxis="y", stackdir="center", dotsize=1, aes(fill=Strain, color=Strain)) + theme_bw() + ggtitle("SCGlu: Area under the curve AUC across FLC concs") + xlab("FLC conc (ug/ml)") + ylab("Area Under the Curve (AUC)")



ggplot(YPGlyresults,aes(x=as.factor(FLC_conc),y=auc_l)) + geom_dotplot(binaxis="y", stackdir="center", dotsize=1, aes(fill=Strain, color=Strain)) + theme_bw() + ggtitle("YPGly: Area under the curve AUC across FLC concs") + xlab("FLC conc (ug/ml)") + ylab("Area Under the Curve (AUC)")

```


SCGlu grows much faster than YPGly and might grow to completion within 24 hours. hence, use only 24 hours of data for re-analysis.

```{r 24hrs}
dataod_24=data_od[data_od$time<=24,]

gcout_24 <- SummarizeGrowthByPlate(dataod_24, 
                                 plot_fit = FALSE) %>% 
  left_join(samplesheet, by = c("sample" = "Well")) %>%
  rename(well = sample, sample = Sample)

gcsampleout_24=gcout_24[!is.na(gcout_24$Strain),]
write.csv(gcsampleout_24,"output24hrs.csv")

SCGlu24=gcsampleout_24[gcsampleout_24$Media=="SCGlu",]

ggplot(SCGlu24,aes(x=as.factor(FLC_conc),y=auc_l)) + geom_dotplot(binaxis="y", stackdir="center", dotsize=1, aes(fill=Strain, color=Strain)) + theme_bw() + ggtitle("SCGlu: Area under the curve AUC across FLC concs, first 24 hrs") + xlab("FLC conc (ug/ml)") + ylab("Area Under the Curve (AUC)")
```